---
title: "R Notebook"
output: html_notebook
---

# Vectors

## Atomic vectors

Atomic vectors are usually created with `c()`:
```{r}
dbl_var <- c(1, 2.5, 4.5)
# With the L suffix, you get an integer rather than a double 
int_var <- c(1L, 6L, 10L)
# Use TRUE and FALSE (or T and F) to create logical vectors 
log_var <- c(TRUE, FALSE, T, F)
chr_var <- c("these are", "some strings")
```

Nested `c()`s are ignored:
```{r}
c(1, c(2, c(3, 4)))
c(1, 2, 3, 4) 
```

### Types and tests

Given a vector, you can determine its type with `typeof()`, or check if it’s a specific type with an `is` function:
```{r}
int_var <- c(1L, 6L, 10L)
typeof(int_var)
is.integer(int_var)
is.atomic(int_var)
is.numeric(int_var)
```

```{r}
dbl_var <- c(1, 2.5, 4.5)
typeof(dbl_var)
is.double(dbl_var)
is.atomic(dbl_var)
is.numeric(dbl_var)
```

### Coersion

```{r}
str(c("a", 1))
```

When a logical vector is coerced to an integer or double, TRUE becomes 1 and FALSE becomes 0
```{r}
x <- c(FALSE, FALSE, TRUE) 
as.numeric(x)
sum(x)
mean(x)
```

Implicit coersion:
```{r}
sin(x)
typeof(sin(x))
```

### Lists

Lists are different from atomic vectors because their elements can be of any type, including lists:
```{r}
x <- list(1:3, "a", c(TRUE, FALSE, TRUE), c(2.3, 5.9)) 
str(x)
```

A list can contain other lists:
```{r}
x <- list(list(list(list()))) 
str(x)
is.recursive(x)
```

A data frame is a list:
```{r}
str(mtcars)
is.list(mtcars)
```

## Attributes

All objects can have arbitrary additional attributes, used to store meta- data about the object. Attributes can be thought of as a named list (with unique names). Attributes can be accessed individually with attr() or all at once (as a list) with attributes().

```{r}
y <- 1:10
str(attributes(y))
attr(y, "my_attribute") <- "This is a vector" 
attr(y, "my_attribute")
str(attributes(y))
```

The `structure()` function returns a new object with modified attributes:
```{r}
structure(1:10, my_attribute = "This is a vector")
```

By default, most attributes are lost when modifying a vector:
```{r}
 attributes(y[1])
 attributes(sum(y))
```

### Names

Three ways to create names:
```{r}
x <- c(a=1,b=2,c=3)
y <- 1:3; names(y) <- c("a", "b", "c")
z <- setNames(1:3, c("a", "b", "c"))
names(z)
```

## Factors

A factor is a vector that can contain only predefined values, and is used to store categorical data. Note that factors are integers, not strings.
```{r}
x <- factor(c("a", "b", "b", "a"))
x
class(x)
levels(x)
typeof(x)
```

Factors are useful when you know the possible values a variable may take, even if you don’t see all values in a given dataset.
```{r}
sex_char <- c("m", "m", "m")
sex_factor <- factor(sex_char, levels = c("m", "f"))
table(sex_factor)
```

# Matrices and Arrays

Matrices and arrays are created with matrix() and array(), or by using the assignment form of dim():
```{r}
a <- matrix(1:6, ncol = 3, nrow = 2)
b <- array(1:12, c(2, 3, 2))
c<-1:6
dim(c) <- c(3, 2) 
c
dim(c) <- c(2, 3)
c
```

`length()` generalizes to `nrow()` and `ncol()` for matrices, and `dim()` for arrays.
```{r}
length(a)
nrow(a)
ncol(a)
length(b)
dim(b)
```

`names()` generalizes to `rownames()` and `colnames()` for matrices, and
`dimnames()`, a list of character vectors, for arrays.
```{r}
rownames(a) <- c("A", "B") 
colnames(a) <- c("a", "b", "c")
a
```

```{r}
dimnames(b) <- list(c("one", "two"), c("a", "b", "c"), c("A", "B"))
b
```


