---
title: "Vocabulary"
output: html_document
date: "2023-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Important operators and assignment

### match

`match` returns a vector of the positions of (first) matches of its first argument in its second.

```{r}
match(c(10, 8, 6, 4, 2), c(9, 2, 3, 5, 7))
```
```{r}
match(c(10, 8, 6, 4, 2, 0), c(7, 5, 3, 2), nomatch=0)
```


`%in%` is a more intuitive interface as a binary operator, which returns a logical vector indicating if there is a match or not for its left operand. %in% is currently defined as 
`"%in%" <- function(x, table) match(x, table, nomatch = 0) > 0`

```{r}
1:10 %in% c(1,3,5,9)
1:10
```
Find all characters in a vector of strings:
```{r}
sstr <- c("c","ab","B","bba","c",NA,"@","bla","a","Ba","%")
sstr[sstr %in% c(letters, LETTERS)]
```

### `head` and `tail`

```{r}
tail(letters, n = 3)
```

```{r}
head(mtcars)
```

```{r}
head(freeny.x,  c(6L, 2L))
```

### `with`

Evaluate an R expression in an environment constructed from data, possibly modifying (a copy of) the original data.

```{r}
with(mtcars, mpg[cyl == 8  &  disp > 350])
```

### `assign` and `get`

Assign a value to a name in an environment.

```{r}
for(i in 1:6) { #-- Create objects  'r.1', 'r.2', ... 'r.6' --
    nam <- paste("r", i, sep = ".")
    assign(nam, 1:i)
}
ls(pattern = "^r..$")
```

```{r}
get("r.3")
```

##Comparison

### `all.equal`

```{r}
all.equal(4*atan(1), 3.14159)
all.equal(4*atan(1), 3.141592654)
```


```{r}
x1 <- x2 <- (1:100)/10;  x2[2] <- 1.1*x1[2]
head(x1)
head(x2)
all.equal(x1, x2)
```

### `identical`

The safe and reliable way to test two objects for being exactly equal. It returns TRUE in this case, FALSE in every other case.

```{r}
identical(1, NULL) ## FALSE -- don't try this with ==
identical(1, 1.)   ## TRUE in R (both are stored as doubles)
identical(1, as.integer(1)) ## FALSE, stored as different types
identical(1, 1L)
identical(1, TRUE)
identical(1, "1")
```

### `is.na`

```{r}
is.na(c(1, NA))
```

### `is.finite`

```{r}
is.finite(c(1/0, 0/1, 1/1, 0/0))
```

## Basic math

### Power

```{r}
3 ^ 2
2 ^ 16
3 ** 3
```

### Division and remainder

```{r}
45 %% 12 # remainder
45 %/% 12 # integer division
```

### `signif`

```{r}
signif(4*atan(1), 5)
```

### Cumulative functions

```{r}
cumsum(1:10)
cumprod(1:10)
cummin(c(3:1, 2:0, 4:2))
cummax(c(3:1, 2:0, 4:2))
```


### `range`

`range` returns a vector containing the minimum and maximum of all the given arguments.

```{r}
range(stats::rnorm(1000))
```

### Statistics

```{r}
mean(stats::rnorm(100))
median(runif(100))
sd(runif(100))
var(runif(100))
```

### `any` and `all`

```{r}
r = rbinom(10, 5, 0.5)
r
all(r)
any(r == 0)
```

### `which`

Give the TRUE indices of a logical object, allowing for array indices.

```{r}
which(LETTERS == "R")
which((1:12)%%2 == 1)
```

## Vectors and matrices

### `c` and `matrix`

```{r}
c(1, 2, 4)
matrix(1:10, nrow=2)
```

### `length`, `dim`, `nrow`, `ncol`

```{r}
m = matrix(-5:2, nrow=4)
length(m)
dim(m)
nrow(m)
ncol(m)
```

### `cbind` and `rbind`

```{r}
rbind(1, 1:7)
cbind(1:3, 4:6)
```

### `t`

```{r}
a <- matrix(1:30, 5, 6)
a
t(a)
```


### `diag`

Extract or replace the diagonal of a matrix, or construct a diagonal matrix.

```{r}
diag(c(1,2,4))
```

```{r}
# diag(<var-cov-matrix>) = variances
corr.matrix = var(M <- cbind(X = 1:10, Y = rnorm(10)))
corr.matrix
diag(corr.matrix)
```

## Making vectors

### `rep` and `rep_len`

`rep` replicates the values in `x`.

```{r}
rep(1:4, 2)
rep(1:4, c(2,1,2,1))
rep_len(letters[1:5], 11)
```

### `seq`

Generate regular sequences.

```{r}
seq(0, 1, length.out = 21)
seq(stats::rnorm(20))
seq(1, 9, by = 2)
seq(0, 10, by = pi)
```

```{r}
seq_len(5)
seq_along(c(5, 9, 10, -1))
```

### `rev`

`rev` provides a reversed version of its argument.

```{r}
rev(1:7)
rev(c(1:5, 5:3))
```

### `sample`

`sample` takes a sample of the specified size from the elements of `x` using either with or without replacement.

```{r}
x <- 1:12
sample(x, 7)
sample(x, 10, replace = TRUE)
```

### `choose`, `factorial`, `combn`

```{r}
choose(4, 2)
factorial(5)
combn(letters[1:5], 3)
```

### is/as

```{r}
as.numeric(c(-1, "2", 3.14, FALSE))
as.character(c(TRUE, "42", 15, 10L))
is.logical(c(1 > 0, FALSE, 12))
```

## Lists and data.frames

## `list` and `unlist`

```{r}
list(1.2, "34", FALSE, 3L, matrix(1:6, 2, 3))
```

Given a list structure `x`, `unlist` simplifies it to produce a vector which contains all the atomic components which occur in `x`.

```{r}
l1 <- list(a = "a", b = 2, c = pi+2i)
unlist(l1) # a character vector
```

### `data.frame`

```{r}
L3 <- LETTERS[1:3]
char <- sample(L3, 10, replace = TRUE)
(d <- data.frame(x = 1, y = 1:10, char = char))
```

```{r}
n = 43
if (n %% 2 == 0) {
  cat("even")
} else {
  cat("odd")
}
```

```{r}
for (i in 1:5) {
  print(i)
  print(-i)
}
```

```{r}
x <- c(6:-4)
sqrt(ifelse(x >= 0, x, NA))
```

```{r}
require(stats)
centre <- function(x, type) {
  switch(type,
         mean = mean(x),
         median = median(x),
         trimmed = mean(x, trim = .1))
}
x <- rcauchy(10)
centre(x, "mean")
centre(x, "median")
centre(x, "trimmed")
```

